---
title: "Getting started with gsm.vizr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with gsm.vizr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(gsm.vizr)
```

## What this package owns

Three layers, with one owner each. **gsm.viz** is the JavaScript library: it
decides how a bar chart looks and behaves, and its documentation is the source
of truth for every spec key. **gsm.vizr** is this package: it vendors one copy
of the gsm.viz bundle, owns the R-to-JavaScript serialization contract, and
ships the Shiny, event and proxy plumbing. **Report packages** — gsm.qtl,
gsm.kri, gsm.endpoints and the rest — build data frames and specs and hand them
here.

The wrapper is deliberately thin. R lists map 1:1 onto the JS spec, so there is
no second vocabulary to learn and no translation layer to fall behind upstream.

## A first chart

`bars()` takes a data frame and a spec. `bars_spec()` builds the spec and
validates it in R, mirroring gsm.viz's own `validateSpec.js` — so a malformed
spec fails here, with the same message, instead of silently in a browser.

```{r basic}
dfSites <- data.frame(
  site = rep(c("S-001", "S-002", "S-003"), times = c(5, 3, 2)),
  flag = rep(c("2", "1", "2", "1", "0"), times = 2)
)

bars(dfSites, bars_spec(x = "site", fill = "flag"))
```

Aesthetics nest under `mapping`; everything else is a top-level spec key:

```{r spec-shape}
str(bars_spec(x = "invid", y = "n", fill = "flag", stat = "identity"))
```

### Factors carry order

Category and stack order come from factor levels automatically, so `forcats`
verbs travel to the browser without a second ordering argument:

```{r factors}
dfOrdered <- data.frame(
  site = factor(c("S-001", "S-002"), levels = c("S-002", "S-001")),
  flag = factor(c("2", "1"), levels = c("2", "1", "0"))
)
spec <- bars_spec(x = "site", fill = "flag")
# scales$x$order and scales$fill$order are filled in from the levels at render.
bars(dfOrdered, spec)
```

An order you set yourself always wins.

## Escaping to JavaScript

Prefer the JSON-safe options — `tooltip$format`, label format strings — where
they suffice. When a slot genuinely needs a function, `js_hook()` marks it:

```{r hook}
bars(
  dfSites,
  bars_spec(
    x = "site", fill = "flag",
    tooltip = list(formatter = js_hook(
      "function (value, context, details) { return 'Subjects: ' + value; }"
    ))
  )
)
```

Hooks are revived in the browser on an allowlist only:
`tooltip$formatter`, `callbacks$onClick`, `callbacks$onHover`,
`callbacks$onSelect`, and the two annotation label formatters. A `js_hook()`
anywhere else is an R error naming where it was found, rather than a string
quietly serialized where a function was expected.

## The event contract

Every chart dispatches a bubbling `gsm-viz-select` DOM event, so a report can
listen once at the document level rather than wiring each chart:

```js
document.addEventListener('gsm-viz-select', function (e) {
  console.log(e.detail);
});
```

Two detail shapes:

| field | click | select |
| --- | --- | --- |
| `type` | `"click"` | `"select"` |
| `chartId` | the widget id | the widget id |
| `category` | clicked category | — |
| `fill` | clicked segment, or `null` | — |
| `datum` | contributing row(s) | — |
| `selection` | — | `{type, values}` |
| `metadata` | your `metadata` list | your `metadata` list |

### `datum` changes shape with `stat`

This is the one field that is not a fixed type, and it is worth reading twice.
gsm.viz sets it to **the single contributing row** under `stat = "identity"`,
and to **the array of aggregated rows** under `stat = "count"`. gsm.vizr
forwards it unchanged.

So `detail.datum.invid` works on a pre-aggregated chart and silently yields
`undefined` on a counting one, where you need `detail.datum[0].invid` or a
reduction across the array. Branch on the stat mode you built the chart with.

## Shiny

`barsOutput()` and `renderBars()` are the usual htmlwidgets pair. Click and
select details also arrive as `input$<id>_click` and `input$<id>_select`.

`bars_proxy()` updates a rendered chart in place, without re-rendering it:

```r
observeEvent(input$refresh, {
  bars_proxy("chart") |>
    proxy_update_data(dfNew) |>
    proxy_select_category("S-001")
})
```

Selection verbs are **silent by default**: a server-driven selection does not
echo a select event back to the server, which is what stops an update loop.
Pass `silent = FALSE` when you do want the round trip.

`proxy_update_data()` without a spec deliberately reuses the browser's live
spec. The chart's own controls (the position and category toggles) mutate that
spec, so the browser copy is authoritative — re-pushing a spec cached in R
would quietly undo whatever the user had toggled.

**In 0.1.0 the proxy verbs apply to `bars()` widgets only.** A verb aimed at a
`facet_bars()` widget raises an error naming the widget and the verb, rather
than doing nothing. Fanning verbs across facet panels needs per-panel semantics
that upstream does not yet define — should selecting a category select it in
one panel or all of them? — so guessing was deferred rather than baked in.

## Driving a chart without Shiny

`bars_proxy()` needs a Shiny session. Plenty of consumers do not have one:
gsm.endpoints' reports and gsm.kri's premature-deaths report are plain
`output: html_document`. Their drilldowns reach exactly the same functions from
report JavaScript, through `chart.helpers` on the rendered element:

```js
var el = document.getElementById('my-chart-id');
var ch = el.gsmChart;

// Same call the updateData proxy verb makes. Passing the live spec preserves
// whatever the chart's own toggles have changed.
ch.helpers.updateData(ch, rows, ch.data._spec_);

ch.helpers.selectCategory(ch, ['S-001'], undefined, { _silent: true });
ch.helpers.clearSelection(ch, undefined, { _silent: true });
ch.helpers.getSelection(ch);
```

Give the chart a stable id through `metadata` so report JavaScript can find it:

```{r static-id, eval = FALSE}
bars(dfSites, bars_spec(x = "site", fill = "flag"),
  metadata = list(chartId = "my-chart-id", level = "site")
)
```

`metadata` is echoed back in every event detail, so it is also where report-level
routing keys belong.

## For package authors

If your package ships its own custom widget and needs `window.gsmViz`, declare
the shared dependency instead of vendoring a second copy of the bundle:

```{r dependency}
dep <- html_dependency_gsm_viz()
c(name = dep$name, version = dep$version)
```

Attach it to your widget's tags with `htmltools::attachDependencies()`. One
vendored bundle then serves every consumer, and upgrading it is one change in
one place.

## Sizing

Two independent knobs, often confused:

- `minHeight` (default 500px) is a **floor**, applied as a CSS minimum so a
  chart cannot collapse.
- `theme$pxPerCategory` (upstream default 30) is the **growth rate** under
  `theme$dynamicSizing`: height scales with the category count.

They compose — set both to reproduce a caller's own sizing curve. A report that
sized figures at 25px per site with a 500px minimum is:

```{r sizing, eval = FALSE}
bars_spec(
  x = "invid", y = "n", stat = "identity", orientation = "horizontal",
  theme = list(dynamicSizing = TRUE, pxPerCategory = 25)
)
```
