---
title: "Chart gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chart gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE
)
```

```{r setup}
library(gsm.vizr)
```

Every chart shape gsm.vizr covers, in the combinations report packages actually
use. The `bars()` shapes below are asserted in the package's Playwright suite,
so that matrix and the tests stay in step.

```{r data}
dfSites <- data.frame(
  site = rep(c("S-001", "S-002", "S-003"), times = c(5, 3, 2)),
  flag = rep(c("2", "1", "2", "1", "0"), times = 2)
)
dfCounts <- data.frame(site = sprintf("S-%02d", 1:8), n = c(9, 3, 7, 1, 5, 2, 8, 4))
dfFlags <- data.frame(
  invid = factor(rep(c("INV-3", "INV-1", "INV-2"), each = 2),
    levels = c("INV-3", "INV-1", "INV-2")
  ),
  flag = factor(rep(c("2", "1"), times = 3), levels = c("2", "1")),
  n = c(5, 2, 4, 1, 2, 2)
)
```

# Generic bar charts

## Vertical stack (count)

The default: one row per observation, gsm.viz does the counting.

```{r basic}
bars(dfSites, bars_spec(x = "site", fill = "flag"))
```

## Horizontal stack, pre-aggregated, named colours, captions

```{r hbar}
bars(dfFlags, bars_spec(
  x = "invid", y = "n", fill = "flag",
  orientation = "horizontal", stat = "identity",
  scales = list(fill = list(colors = c(`2` = "#c91b1d", `1` = "#fdb84c"))),
  labels = list(captions = c("Footnote line 1.", "Footnote line 2."))
))
```

## Position: dodge

```{r dodge}
bars(dfSites, bars_spec(x = "site", fill = "flag", position = "dodge"))
```

## 100% normalized

`position = "fill"` and `stat = "percent"` produce the same chart — gsm.viz
normalizes the former into the latter.

```{r fill}
bars(dfSites, bars_spec(x = "site", fill = "flag", position = "fill"))
```

## Single series, frequency sort, and Top-N

```{r identity}
bars(dfCounts, bars_spec(x = "site", y = "n", stat = "identity"))
```

```{r sort}
bars(dfCounts, bars_spec(
  x = "site", y = "n", stat = "identity",
  scales = list(x = list(sort = "total", sortDir = "desc"))
))
```

```{r topn}
bars(dfCounts, bars_spec(x = "site", y = "n", stat = "identity", nCategories = 3))
```

## Labels, tooltips, reference lines

```{r seg-labels}
bars(dfFlags, bars_spec(
  x = "invid", y = "n", fill = "flag", stat = "identity",
  annotations = list(labels = list(
    segment = list(display = TRUE),
    total = list(display = TRUE)
  ))
))
```

```{r reflines}
bars(dfCounts, bars_spec(
  x = "site", y = "n", stat = "identity",
  annotations = list(referenceLines = list(
    list(value = 5, label = "Threshold", color = "#c91b1d")
  ))
))
```

## Zoom, dense legend, multi-select

```{r zoom}
bars(dfCounts, bars_spec(
  x = "site", y = "n", stat = "identity",
  zoom = list(enabled = TRUE)
))
```

```{r multi-select}
bars(dfSites, bars_spec(
  x = "site", fill = "flag",
  selection = list(enabled = TRUE, multiple = TRUE)
))
```

## Facet small multiples

```{r facet}
dfFacet <- data.frame(
  site = rep(c("S-001", "S-002"), 4),
  flag = rep(c("2", "1"), each = 4),
  country = rep(c("USA", "CAN"), each = 4)
)
facet_bars(dfFacet, bars_spec(x = "site", fill = "flag"),
  facet_spec("country", order = c("USA", "CAN"), nCol = 2)
)
```

## Downstream replicas

The shapes the migrating report packages actually render.

### Participant-level counts, normalized, with footnotes

```{r qtl-replica}
set.seed(158)
dfSubj <- data.frame(
  invid = factor(sample(sprintf("INV-%d", 1:6), 60, replace = TRUE)),
  category = factor(
    sample(c("Eligible", "Excluded", "Pending"), 60, replace = TRUE),
    levels = c("Eligible", "Excluded", "Pending")
  )
)
bars(dfSubj, bars_spec(
  x = "invid", fill = "category", orientation = "horizontal",
  position = "fill",
  theme = list(dynamicSizing = TRUE),
  labels = list(captions = c(
    "Counts of participants by investigator.",
    "Source: synthetic gallery data."
  ))
))
```

### Pre-aggregated eligibility stack, with titles

The dominant real shape: already summarized, horizontal, plain stack, with a
chart title and both axis titles.

```{r qtl-identity}
dfElig <- data.frame(
  invid = factor(rep(sprintf("INV-%d", 1:4), each = 3),
    levels = sprintf("INV-%d", 4:1)
  ),
  criteria = factor(rep(c("Ineligible", "No Eligibility Risk", "Neither"), 4),
    levels = c("Ineligible", "No Eligibility Risk", "Neither")
  ),
  n = c(3, 8, 1, 5, 6, 2, 1, 9, 3, 4, 4, 4)
)
bars(dfElig, bars_spec(
  x = "invid", y = "n", fill = "criteria", stat = "identity",
  orientation = "horizontal",
  scales = list(
    x = list(label = "Investigator"),
    y = list(label = "Participants"),
    fill = list(colors = c(
      Ineligible = "#FF5859", `No Eligibility Risk` = "#00BFC4",
      Neither = "#7CAE00"
    ))
  ),
  labels = list(
    title = "Eligibility criteria by investigator",
    captions = "Participants screened between 2026-01-01 and 2026-06-30."
  )
))
```

### Endpoint status with a two-line tooltip

```{r endpoints-replica}
dfPfs <- data.frame(
  site = rep(c("S-001", "S-002", "S-003"), each = 2),
  status = factor(rep(c("Event", "Censored"), times = 3),
    levels = c("Event", "Censored")
  ),
  n = c(12, 30, 7, 25, 15, 18),
  pct = c(28.6, 71.4, 21.9, 78.1, 45.5, 54.5)
)
bars(dfPfs, bars_spec(
  x = "site", y = "n", fill = "status", stat = "identity",
  scales = list(fill = list(colors = c(Event = "#c91b1d", Censored = "#4e79a7"))),
  annotations = list(labels = list(segment = list(display = TRUE))),
  tooltip = list(formatter = js_hook(
    "function (value, context, details) {
       var d = (details && details.datum) || {};
       return [d.status + ': ' + d.n, 'Rate: ' + d.pct + '%'];
     }"
  ))
))
```

## Sizing at report scale

24 investigators, growing at 25px each above the 500px floor — the sizing curve
a report package would otherwise compute itself.

```{r dynamic-many}
set.seed(24)
dfMany <- data.frame(
  invid = factor(rep(sprintf("INV-%02d", 1:24), each = 2)),
  flag = factor(rep(c("2", "1"), 24), levels = c("2", "1")),
  n = sample(1:9, 48, replace = TRUE)
)
bars(dfMany, bars_spec(
  x = "invid", y = "n", fill = "flag", stat = "identity",
  orientation = "horizontal",
  theme = list(dynamicSizing = TRUE, pxPerCategory = 25)
))
```

## Empty data

Renders without error, though without a placeholder message.

```{r empty}
bars(
  data.frame(site = character(), flag = character()),
  bars_spec(x = "site", fill = "flag")
)
```

# Legacy KRI widgets

Relocated from gsm.kri, these wrappers drive gsm.viz's earlier KRI renderers —
`barChart`, `scatterPlot`, `timeSeries`, `groupOverview` — from the same
vendored bundle the charts above use. They take the gsm reporting data frames
directly rather than a `bars_spec()`, and build their configuration internally
through `MakeChartConfig()`. They exist so KRI reports keep working across the
move; new charts should reach for `bars()`.

```{r kri-data}
dfKri <- gsm.core::reportingResults %>%
  dplyr::filter(MetricID == "Analysis_kri0001")
dfKriLatest <- dfKri %>% dplyr::filter(SnapshotDate == max(SnapshotDate))
lKriMetric <- gsm.core::reportingMetrics %>%
  dplyr::filter(MetricID == "Analysis_kri0001") %>%
  as.list()
dfKriBounds <- gsm.core::reportingBounds %>%
  dplyr::filter(
    MetricID == "Analysis_kri0001" & SnapshotDate == max(SnapshotDate)
  )
```

## Bar chart

Group-level results for one metric at the latest snapshot, with the outcome
dropdown and site highlight control.

```{r widget-barchart}
Widget_BarChart(
  dfResults = dfKriLatest,
  dfGroups = gsm.core::reportingGroups,
  lMetric = lKriMetric,
  vThreshold = lKriMetric$Threshold
)
```

## Scatter plot

Denominator against numerator, with the metric's bounds behind the points.

```{r widget-scatterplot}
Widget_ScatterPlot(
  dfResults = dfKriLatest,
  dfGroups = gsm.core::reportingGroups,
  lMetric = lKriMetric,
  dfBounds = dfKriBounds
)
```

## Time series

The same metric across every snapshot rather than only the latest.

```{r widget-timeseries}
Widget_TimeSeries(
  dfResults = dfKri,
  dfGroups = gsm.core::reportingGroups,
  lMetric = lKriMetric,
  vThreshold = lKriMetric$Threshold
)
```

## Group overview

A table rather than a chart, and the one widget that spans every metric at once.
The default subset keeps the groups carrying at least one red flag.

```{r widget-groupoverview}
Widget_GroupOverview(
  dfResults = dplyr::filter(
    gsm.core::reportingResults,
    SnapshotDate == max(SnapshotDate)
  ),
  dfMetrics = gsm.core::reportingMetrics,
  dfGroups = gsm.core::reportingGroups
)
```
